Evaluating FAIR Digital Object as a distributed object system

### 2023-03-21 latex/table4.tex
updated bibliography entries

\caption{Comparing FAIR Digital Object (with the DOIP 2.0 protocol \autocite{13TcbsZF6}) and Web technologies (using Linked Data) as middleware infrastructures \autocite{hRzcHhPD}

@@

  & DOIP types nested with parents, implying shared FDO structures (unclear if operations are inherited). FDO establishes need for multiple Data Type Registries (e.g.~managed by a community for a particular domain). Semantics of type system currently undefined for FDO and DOIP, syntactic types can also piggyback of FDO type's schema (e.g.~\href{(https://www.cordra.org/documentation/design/schemas.html\#schema-references)}{CORDRA \texttt{\$ref}} use of \href{https://json-schema.org/draft/2020-12/json-schema-core.html\#references}{JSON Schema references} \autocite{Draftbhuttonjsonschema00})
  & Syntactically Media Type with multiple suffixes \autocite{Draftietfmediamansuffixes00MediaTypes} (mainly used with \texttt{+json}), declaration of subtypes as profiles (RFC6906) \autocite{rfc6906}. In metadata, semantic type systems (RDFS \autocite{w3-rdf-schema}), OWL2 \autocite{w3-owl2-overview}, SKOS \autocite{w3-skos-primer}). OpenAPI 3 \autocite{OpenAPISpecificationV3} \href{https://spec.openapis.org/oas/v3.1.0\#composition-and-inheritance-polymorphism}{inheritance and Polymorphism}. XML \texttt{xsd:schemaLocation} or \texttt{xsd:type} \autocite{w3-xmlschema11}, JSON \texttt{\$schema} \autocite{Draftbhuttonjsonschema00}), JSON-LD \texttt{@context} \autocite{w3-json-ld}. Large number of domain-specific and general ontologies define semantic types, but finding and selecting remains a challenge. \\

@@

  & CRUD predefined in \href{https://datatracker.ietf.org/doc/html/rfc7231\#section-4.3}{HTTP methods} \autocite{HE7Ikwwl}, (\href{https://www.iana.org/assignments/http-methods/http-methods.xhtml}{extended by registration}), URI Templates \autocite{rfc6570}, \href{https://spec.openapis.org/oas/v3.1.0.html\#operation-object}{OpenAPI operations} \autocite{OpenAPISpecificationV3}, HATEOAS\footnote{HATEOAS: Hypermedia as the Engine of Application State \autocite{fieldingArchitecturalStylesDesign2000a}, an important element of the REST architectural style.} incl.~Hydra \autocite{HydraW3CCommunity}, schema.org Actions \autocite{SchemaOrgActions}, JSON HAL \autocite{Draftkellyjsonhal08}) \& Link headers (RFC8288) \autocite{rfc8288} \\

>>>>

\caption{Comparing FAIR Digital Object (with the DOIP 2.0 protocol \autocite{foundationDigitalObjectInterface}) and Web technologies (using Linked Data) as middleware infrastructures \autocite{zarrasComparisonFrameworkMiddleware2004a}

@@

  & DOIP types nested with parents, implying shared FDO structures (unclear if operations are inherited). FDO establishes need for multiple Data Type Registries (e.g.~managed by a community for a particular domain). Semantics of type system currently undefined for FDO and DOIP, syntactic types can also piggyback of FDO type's schema (e.g.~\href{(https://www.cordra.org/documentation/design/schemas.html\#schema-references)}{CORDRA \texttt{\$ref}} use of \href{https://json-schema.org/draft/2020-12/json-schema-core.html\#references}{JSON Schema references} \autocite{Draftbhuttonjsonschema})
  & Syntactically Media Type with multiple suffixes \autocite{Draftietfmediamansuffixes00MediaTypes} (mainly used with \texttt{+json}), declaration of subtypes as profiles (RFC6906) \autocite{rfc6906}. In metadata, semantic type systems (RDFS \autocite{w3-rdf-schema}), OWL2 \autocite{w3-owl2-overview}, SKOS \autocite{w3-skos-primer}). OpenAPI 3 \autocite{OpenAPISpecificationV3} \href{https://spec.openapis.org/oas/v3.1.0\#composition-and-inheritance-polymorphism}{inheritance and Polymorphism}. XML \texttt{xsd:schemaLocation} or \texttt{xsd:type} \autocite{w3-xmlschema11}, JSON \texttt{\$schema} \autocite{Draftbhuttonjsonschema}), JSON-LD \texttt{@context} \autocite{w3-json-ld}. Large number of domain-specific and general ontologies define semantic types, but finding and selecting remains a challenge. \\

@@

  & CRUD predefined in \href{https://datatracker.ietf.org/doc/html/rfc7231\#section-4.3}{HTTP methods} \autocite{rfc7231}, (\href{https://www.iana.org/assignments/http-methods/http-methods.xhtml}{extended by registration}), URI Templates \autocite{rfc6570}, \href{https://spec.openapis.org/oas/v3.1.0.html\#operation-object}{OpenAPI operations} \autocite{OpenAPISpecificationV3}, HATEOAS\footnote{HATEOAS: Hypermedia as the Engine of Application State \autocite{fieldingArchitecturalStylesDesign2000a}, an important element of the REST architectural style.} incl.~Hydra \autocite{HydraW3CCommunity}, schema.org Actions \autocite{SchemaOrgActions}, JSON HAL \autocite{Draftkellyjsonhal08}) \& Link headers (RFC8288) \autocite{rfc8288} \\

### 2023-03-21 latex/table5.tex
updated bibliography entries

\caption{Assessing RDA's FAIR Data Maturity Model \autocite{1GrKTFaK2,UzQhqk0M} (first 2 columns) against the FDO guidelines \autocite{RwvirqWg}, FDO implemented with the protocol DOIPv2 \autocite{13TcbsZF6}, Linked Data Platform (LDP) \autocite{7szz7dwO} and examples from Linked Data practices in general. (--- indicates \emph{Unspecified}, may be possible with additional conventions)

@@

  & Large number of general and domain-specific vocabularies can make it hard to find relevant attributes. Rough consensus on kernel metadata: schema.org \autocite{10pzKCEUE}, Dublin Core Terms \autocite{DCMIMetadataTerms}, DCAT \autocite{w3-vocab-dcat-2}, FOAF \autocite{FOAFVocabularySpecification} \\

>>>>

\caption{Assessing RDA's FAIR Data Maturity Model \autocite{groupFAIRDataMaturity2020,bahimFAIRDataMaturity2020a} (first 2 columns) against the FDO guidelines \autocite{boninoFAIRDigitalObject}, FDO implemented with the protocol DOIPv2 \autocite{foundationDigitalObjectInterface}, Linked Data Platform (LDP) \autocite{FDOFramework} and examples from Linked Data practices in general. (--- indicates \emph{Unspecified}, may be possible with additional conventions)

@@

  & Large number of general and domain-specific vocabularies can make it hard to find relevant attributes. Rough consensus on kernel metadata: schema.org \autocite{SchemaOrg}, Dublin Core Terms \autocite{DCMIMetadataTerms}, DCAT \autocite{w3-vocab-dcat-2}, FOAF \autocite{FOAFVocabularySpecification} \\
